Preliminary Draft

### 2021-07-21 article/draft.tex
minor changes

During the COVID-19 global health pandemic platforms have upgraded their guidelines to include a set of rules to tackle the propagation of potentially harmful content (see table \ref{summary} in Appendix \ref{links}). Those policies are enforced via existing actions used by the platforms to tackle other rules' violations, such as: labelling content to provide more context or indicate falsehood, publishing a list of terms or topics that will be flagged, suspending accounts, implementing strike systems, reducing the visibility of content, etc. As each platform is a private company, those {\it new} policies are not coordinated and are implemented in different ways across platforms. Such targeted policies show the willingness of mainstream platforms to enhance the quality of the online conversation, but also sheds light on the lack of specific policies to tackle misinformation in general. In particular, policies regarding misinformation are not part of the set of platform rules or community guidelines (as of July 2021). The 2019 report of the Facebook Data Transparency Advisory Group (DTAG) states that ``{\it DTAG was not tasked with evaluating any of the following: (...) Facebook’s policies with respect to “fake news” or misinformation, as neither of these categories were counted as violations within the first two versions of the Community Standards Enforcement Report.}''. %Facebook's strategy to tackle ``False News'' is three fold : Remove, Reduce, Inform. It is explained via a blog post on the facebook Newsroom. Similarly, Twitter communicates about actions related to misinformation via their \href{https://blog.twitter.com/en\_us/authors.TwitterSafety}{Twitter Safety} Blog/account. In a post on YouTube Official Blog, the platform explained its ``Four Rs of Responsibility" and how it raises authoritative content, reduces borderline content and harmful misinformation (see table \ref{summary} in Appendix \ref{links}).

@@

Nevertheless, misinformation specific interventions by mainstream platforms are hard to monitor, study or verify by third parties (e.g. academic community, data journalists, NGOs), as in many cases misleading or false content does not qualify as a violation of a given platform's rules and it does not explicitly appear as a separate category in available transparency reports. This makes the study of online misinformation, the assessment of the impact of platforms' actions to tackle misinformation and their relevancy a burdensome task for the academic community. Hence, in the present article we explain how to verify with data mining mainstream platforms' current actions regarding content with low credibility or false information. We do so by providing a series of examples for different interventions and platforms. We chose to focus on three platforms: Facebook, Twitter and YouTube. Both Facebook and YouTube are in the top three most popular social media platforms in terms of number of users.\footnote{See for example the ranking of the most popular social networks as of April 2021 on Statista: \href{https://www.statista.com/statistics/272014/global-social-networks-ranked-by-number-of-users/}{https://www.statista.com/statistics/272014/global-social-networks-ranked-by-number-of-users/}.} We further choose Twitter because it is a social networking platform with the most news-focused users, according to the Pew Research center (2019)~\cite{pew1}. To collect data from these three platforms, we either used the APIs (Application Programming Interfaces), or web scraping (i.e. retro-engineering the HMTL code of a web page to extract meaningful data, see Table \ref{tab1} for a summary). Minet, a webmining tool developed by the médialab \cite{minet}, was often used, and we scripted our own data mining code when it was necessary.

@@

In this section, we turn to the channel's temporary or permanent suspension policy of YouTube. Whenever a channel publishes a video that violates the community guidelines for the first time they will  usually receive a warning and the content will be removed. For the second time the channel will start receiving strikes. A first strike results in limiting the access of the YouTube channel  for one week, like uploading videos, streaming and other activities. Then a second strike is similar but the suspension will be for two weeks. A third strike results in the termination of the channel. The strike count of a channel lasts 90 days.  In the special case, where a video is in extreme violation of the guidelines, the publishing channel may get terminated without a warning. To illustrate the implementation of this policy we provide two examples for the temporary suspension of the following two YouTube channels: \href{https://www.youtube.com/channel/UCNbIDJNNgaRrXOD7VllIMRQ}{One America news Network} and \href{https://www.youtube.com/user/TonyHeller1}{Tony Heller}.

\begin{figure}[h]
	\centering
	%\begin{multicols}{1}
		\includegraphics[scale=0.21]{./img/oann/fig3_oann.png}
	%\end{multicols}
	%\begin{multicols}{1}
		\includegraphics[scale=0.31]{./img/tony/fig3_tony.png}
	%\end{multicols}

	\caption{Left Panel: Tweet announcing moving to rumble by OANN (Twitter), Twitter ID \href{https://twitter.com/OANN/status/1372238828425998336}{1372238828425998336}. Right Panel: Tony Heller's tweet after getting suspended from YouTube, Twitter ID \href{https://twitter.com/Tony\_Heller/status/1310703852769796097}{1310703852769796097}.}
	\label{fig2_oann}
\end{figure}

First, we investigate the temporary suspension case of the YouTube channel of {\it One America News channel}. This channel received a first strike on November 24, 2020 for the promotion of  a false cure for COVID-19.\footnote{See nbcnews,  \href{https://www.nbcnews.com/news/all/youtube-suspends-oann-violating-its-covid-19-policy-n1248845}{YouTube suspends OANN for violating its Covid-19 policy} nbcnews, Ahiza García-Hodges, 24 11 2020.} We collected the activity of the channel OANN (video counts, view counts) using  the YouTube API v3, between November 2020 and January 2021. For the video counts, we used the playlist endpoint to retrieve the videos uploaded with their publishing date and for the view count we used the IDs of the videos we had from the playlists and via the videos endpoint we retrieved the view counts on June 2021.%\footnote{See the Google documentation  \href{https://developers.google.com/youtube/v3/docs/videos/list}{https://developers.google.com/youtube/v3/docs/videos/list} and \href{https://developers.google.com/youtube/v3/docs/playlists/list}{https://developers.google.com/youtube/v3/docs/playlists/list}}

@@

We further examine the placement of interstitials that indicate a possibly sensitive content. We find that only $9344$ ($2.97\%$) tweets out of $323$ $938$ containing a link marked as False, have an interstitial {\it potentially sensitive content}. To check whether the interstitial {\it potentially sensitive content} is automated or not, we pick out one tweet having shared a link marked as False\footnote{The link marked as False redirects to the article {\it 5G Technology and induction of coronavirus in skin cells – US National Library of Medicine (what David Icke has been saying for months) } and can be found \href{https://davidicke.com/2020/07/22/5g-technology-and-induction-of-coronavirus-in-skin-cells-us-national-library-of-medicine-what-david-icke-has-been-saying-for-months/}{here}.}, which also contains the interstitial {\it potentially sensitive content}. Then we look in the set of $323$ $938$ tweets for other tweets who have shared the exact same link. We find a total of $32$ tweets. Among those $32$ tweets only $5$ tweets had the interstitial ``potentially sensitive content''. Again this points towards the non-automation of interstitials placement. Furthermore, notice that the appearance of interstitials may depend on the settings of a user's account and country or language specific regulations. In particular, in the settings of Twitter account, one can deactivate the display of interstitials for {\it sensitive content}, by ticking the box {\it Display media that may contain sensitive content} in the section {\it content you see}.

@@

We classified the list of $171$ YouTube videos by topic based on its content as shown in table \ref{tab2}. We recorded the number of videos containing an information panel within each category. For COVID-19 and COVID-19 vaccine related videos, more than half of the videos contained an information panel. Nevertheless we noticed that among a set of duplicates of the exact same YouTube video uploaded with different video titles, some contained an information panel while others did not (see figure \ref{duplicates_yt}).  
In addition, we noticed for COVID-19 related videos, when the video title did not include keywords like (Testing, Pandemic, COVID-19, coronavirus), the video might not contain a panel associated with it, and in some cases when the video title contains variations of word COVID like (C O V I D 19 or Cv19) it wouldn’t include a panel either. Therefore, we suspect that YouTube is automatically adding panels to the videos based on the video title and not the content of the video.

\begin{figure}[h]
	\centering
	%\begin{multicols}{1}
		\includegraphics[scale=0.2]{./img/youtube_panels/graph_no_banner.png} 
	%\end{multicols}
	%\begin{multicols}{1}
		\includegraphics[scale=0.2]{./img/youtube_panels/news_without_banner_2.png} 
	%\end{multicols}
	%\begin{multicols}{1}
		\includegraphics[scale=0.2]{./img/youtube_panels/graph_with_banner.png}
	%\end{multicols}
	%\begin{multicols}{1}
		\includegraphics[scale=0.2]{./img/youtube_panels/news_with_banner.png} 
	%\end{multicols}
	\caption{Screenshots taken on July 21, 2021. The screenshots show two duplicates of the same YouTube video uploaded under different titles. Left panel: contains no information panel. Right panel: contains an information panel. }
	\label{duplicates_yt}
\end{figure}

@@

For climate change we found that most of the videos (13 out of 18) did not display an information panel. Lastly, no information panels were found in the general health category set, that included videos that were not related to COVID-19 but contained misinformation concerning cures for cancer, abortion, and viruses. Therefore, it is likely that YouTube add information panels below videos concerning controversial topics that can have misleading information like the climate change, COVID-19, flat earth and vaccine.

>>>>

During the COVID-19 global health pandemic platforms have upgraded their guidelines to include a set of rules to tackle the propagation of potentially harmful content (see table \ref{summary} in Appendix \ref{links}). Those policies are enforced via existing actions used by the platforms to tackle other rules' violations, such as: labelling content to provide more context or indicate falsehood, publishing a list of terms or topics that will be flagged, suspending accounts, implementing strike systems, reducing the visibility of content, etc. As each platform is a private company, those {\it new} policies are not coordinated and are implemented in different ways across platforms. Such targeted policies show the willingness of mainstream platforms to enhance the quality of the online conversation, but also sheds light on the lack of specific policies to tackle misinformation in general. In particular, policies regarding misinformation are not part of the set of platform rules or community guidelines (as of July 2021). The 2019 report of the Facebook Data Transparency Advisory Group (DTAG) states that ``{\it DTAG was not tasked with evaluating any of the following: (...) Facebook’s policies with respect to “fake news” or misinformation, as neither of these categories were counted as violations within the first two versions of the Community Standards Enforcement Report}''. %Facebook's strategy to tackle ``False News'' is three fold : Remove, Reduce, Inform. It is explained via a blog post on the facebook Newsroom. Similarly, Twitter communicates about actions related to misinformation via their \href{https://blog.twitter.com/en\_us/authors.TwitterSafety}{Twitter Safety} Blog/account. In a post on YouTube Official Blog, the platform explained its ``Four Rs of Responsibility" and how it raises authoritative content, reduces borderline content and harmful misinformation (see table \ref{summary} in Appendix \ref{links}).

@@

Misinformation specific interventions by mainstream platforms are hard to monitor, study or verify by third parties (e.g. academic community, data journalists, NGOs), as in many cases misleading or false content does not qualify as a violation of a given platform's rules and it does not explicitly appear as a separate category in available transparency reports. This makes the study of online misinformation, the assessment of the impact of platforms' actions to tackle misinformation and their relevancy a burdensome task for the academic community. Hence, in the present article we explain how to verify with data mining mainstream platforms' current actions regarding content with low credibility or false information. We do so by providing a series of examples for different interventions and platforms. We chose to focus on three platforms: Facebook, Twitter and YouTube. Both Facebook and YouTube are in the top three most popular social media platforms in terms of number of users.\footnote{See for example the ranking of the most popular social networks as of April 2021 on Statista: \href{https://www.statista.com/statistics/272014/global-social-networks-ranked-by-number-of-users/}{https://www.statista.com/statistics/272014/global-social-networks-ranked-by-number-of-users/}.} We further choose Twitter because it is a social networking platform with the most news-focused users, according to the Pew Research center (2019)~\cite{pew1}. To collect data from these three platforms, we either used the APIs (Application Programming Interfaces), or web scraping (i.e. retro-engineering the HMTL code of a web page to extract meaningful data, see Table \ref{tab1} for a summary). Minet, a webmining tool developed by the médialab \cite{minet}, was often used, and we scripted our own data mining code when it was necessary.

@@

In this section, we turn to the channel's temporary or permanent suspension policy of YouTube. Whenever a channel publishes a video that violates the community guidelines for the first time they will  usually receive a warning and the content will be removed. For the second time the channel will start receiving strikes. A first strike results in limiting the access of the YouTube channel  for one week, like uploading videos, streaming and other activities. Then a second strike is similar but the suspension will be for two weeks. A third strike results in the termination of the channel. The strike count of a channel lasts 90 days.  In the special case, where a video is in extreme violation of the guidelines, the publishing channel may get terminated without a warning.

To illustrate the implementation of this policy we provide two examples for the temporary suspension of the following two YouTube channels: \href{https://www.youtube.com/channel/UCNbIDJNNgaRrXOD7VllIMRQ}{One America news Network} and \href{https://www.youtube.com/user/TonyHeller1}{Tony Heller}. The website of One America News Network has a ``low'' factual reporting score according to iffy.news.\footnote{See \href{https://mediabiasfactcheck.com/one-america-news-network/}{mediabiasfactcheck.com/one-america-news-network/}.}  Tony Heller posts regularly blog posts on the website {\it realclimatescience.com}, which also has a ``low'' factual reporting score according to iffy.news.\footnote{See \href{https://mediabiasfactcheck.com/real-climate-science/}{mediabiasfactcheck.com/real-climate-science/}} Both are active on YouTube and both have communicated via their Twitter account about restrictions applied by YouTube over their content (see figure \ref{fig2_oann}).

\begin{figure}[h]
	\centering
	%\begin{multicols}{1}
		\includegraphics[scale=0.21]{./img/oann/fig3_oann.png}
	%\end{multicols}
	%\begin{multicols}{1}
		\includegraphics[scale=0.31]{./img/tony/fig3_tony.png}
	%\end{multicols}
	\caption{Left Panel: Tweet announcing moving to rumble by OANN (Twitter), Twitter ID \href{https://twitter.com/OANN/status/1372238828425998336}{1372238828425998336}. Right Panel: Tony Heller's tweet after getting suspended from YouTube, Twitter ID \href{https://twitter.com/Tony\_Heller/status/1310703852769796097}{1310703852769796097}.}
	\label{fig2_oann}
\end{figure}

First, we investigate the temporary suspension of the YouTube channel of {\it One America News channel}. This channel received a first strike on November 24, 2020 for the promotion of  a false cure for COVID-19 according to the News outlet NBCnews (2020)~\cite{nbcnews}. We collected the activity of the channel OANN (video counts, view counts) using  the YouTube API v3, between November $2020$ and January $2021$. For the video counts, we used the playlist endpoint to retrieve the videos uploaded with their publishing date and for the view count we used the IDs of the videos we had from the playlists and via the videos endpoint we retrieved the view counts on June $2021$.%\footnote{See the Google documentation  \href{https://developers.google.com/youtube/v3/docs/videos/list}{https://developers.google.com/youtube/v3/docs/videos/list} and \href{https://developers.google.com/youtube/v3/docs/playlists/list}{https://developers.google.com/youtube/v3/docs/playlists/list}}

@@

We further examine the placement of interstitials that indicate a possibly sensitive content. We find that only $9344$ ($2.97\%$) tweets out of $323$ $938$ containing a link marked as False, have an interstitial {\it potentially sensitive content}. To check whether the interstitial {\it potentially sensitive content} is automated or not, we pick out one tweet having shared a link marked as False\footnote{The link marked as False redirects to the article {\it 5G Technology and induction of coronavirus in skin cells – US National Library of Medicine (what David Icke has been saying for months) } and can be found \href{https://davidicke.com/2020/07/22/5g-technology-and-induction-of-coronavirus-in-skin-cells-us-national-library-of-medicine-what-david-icke-has-been-saying-for-months/}{here}.}, which also contains the interstitial {\it potentially sensitive content} (see figure \ref{fig_notice}). Then we look in the set of $323$ $938$ tweets for other tweets who have shared the exact same link. We find a total of $32$ tweets. Among those $32$ tweets only $5$ tweets had the interstitial ``potentially sensitive content''. Again this points towards the non-automation of interstitials placement. Furthermore, notice that the appearance of interstitials may depend on the settings of a user's account and country or language specific regulations. In particular, in the settings of Twitter account, one can deactivate the display of interstitials for {\it sensitive content}, by ticking the box {\it Display media that may contain sensitive content} in the section {\it content you see}.

@@

We classified the list of $171$ YouTube videos by topic based on its content as shown in table \ref{tab2}. We recorded the number of videos containing an information panel within each category. For COVID-19 and COVID-19 vaccine related videos, more than half of the videos contained an information panel. Nevertheless we noticed that among a set of duplicates of the exact same YouTube video uploaded with different video titles, some contained an information panel while others did not (see figure \ref{duplicates_yt} for an example).  
In addition, we noticed for COVID-19 related videos, when the video title did not include keywords like (Testing, Pandemic, COVID-19, coronavirus), the video might not contain a panel associated with it, and in some cases when the video title contains variations of word COVID like (C O V I D 19 or Cv19) it wouldn’t include a panel either. Therefore, we suspect that YouTube is automatically adding panels to the videos based on the video title and not the content of the video.

\begin{figure}[h]
	\centering
	%\begin{multicols}{1}
		\includegraphics[scale=0.2]{./img/youtube_panels/graph_with_banner.png} 
	%\end{multicols}
	%\begin{multicols}{1}
		\includegraphics[scale=0.2]{./img/youtube_panels/news_with_banner.png}
	%\end{multicols}
	%\begin{multicols}{1}
		\includegraphics[scale=0.2]{./img/youtube_panels/graph_no_banner.png} 
	%\end{multicols}
	%\begin{multicols}{1}
		\includegraphics[scale=0.2]{./img/youtube_panels/news_without_banner.png} 
	%\end{multicols}
	\caption{Screenshots taken on July 21, 2021. The screenshots show two duplicates of the same YouTube video uploaded under different titles. Top panels: a YouTube video about COVID-19 which contains an information panel, see \href{https://www.youtube.com/watch?v=d9GbVZOcT18}{https://www.youtube.com/watch?v=d9GbVZOcT18}. Bottom panels: a duplicate of the YouTube video in the top panels uploaded under a different title and it does not contain an information panel, see \href{https://www.youtube.com/watch?v=A4RvrEKoNxc}{https://www.youtube.com/watch?v=A4RvrEKoNxc}. }
	\label{duplicates_yt}
\end{figure}

@@

For climate change we found that most of the videos ($13$ out of $18$) did not display an information panel. Lastly, no information panels were found in the general health category set, that included videos that were not related to COVID-19 but contained misinformation concerning cures for cancer, abortion, and viruses. Therefore, it is likely that YouTube add information panels below videos concerning controversial topics that can have misleading information like the climate change, COVID-19, flat earth and vaccine. It appears as if YouTube is focusing its efforts on the current ``hot'' topics with a wide public impact (elections, COVID), leaving room for misinformation to spread concerning other topics.
